---
title: "R-notebook for statistics for the scientific paper Respiratory variation in mitral and tricuspid inflow velocity using real-time phase-contrast cardiovascular magnetic resonance - normal values and reproducibility

pat is the dataset with patient data
io is the dataset with inter-observer data
trt is the dataset with test-retest data
column_headers contains descriptions of the column headers in the three datasets
"

output: html_notebook
---

Install dependencies

```{r}
install.packages("blandr")
install.packages("irr")

library(irr)
library(blandr)
```
```{r}
mean(pat$mit_var_def_1)
sd(pat$mit_var_def_1)
mean(pat$tri_var_def_1)
sd(pat$tri_var_def_1)
```
Normality checks for patients, results indicate that they are normally distributed
```{r}
shapiro.test(pat$mit_var_def_1)
shapiro.test(pat$tri_var_def_1)
```
```{r}
mean(trt$s1mvar)
sd(trt$s1mvar)
mean(trt$s1tvar)
sd(trt$s1tvar)
```
Normality checks for test-retest and inter-observer, results indicate that they are normally distributed
```{r}
shapiro.test(trt$s1mvar)
shapiro.test(trt$s1tvar)
shapiro.test(io$o1mvar)
shapiro.test(io$o1tvar)
```


T-test to check if there is any difference between patient and test-retest session 1 results
```{r}
t.test(pat$mit_var_def_1,trt$s1mvar)
t.test(pat$tri_var_def_1,trt$s1tvar)
```
Calculations to get differences and means to use in the Bland-Altman plots

```{r}
trt$mvar_diff <- trt$s1mvar-trt$s2mvar
trt$mvar_mean <- (trt$s1mvar+trt$s2mvar)/2

trt$tvar_diff <- trt$s1tvar-trt$s2tvar
trt$tvar_mean <- (trt$s1tvar+trt$s2tvar)/2

io$mvar_diff <- io$o1mvar-io$o2mvar
io$mvar_mean <- (io$o1mvar+io$o2mvar)/2

io$tvar_diff <- io$o1tvar-io$o2tvar
io$tvar_mean <- (io$o1tvar+io$o2tvar)/2
```

Normality tests for test-retest and means and SD of the difference
```{r}

shapiro.test(trt$mvar_diff)
shapiro.test(trt$tvar_diff)

mean(trt$mvar_diff)
sd(trt$mvar_diff)

mean(trt$tvar_diff)
sd(trt$tvar_diff)

```

```{r}

shapiro.test(io$mvar_diff)
shapiro.test(io$tvar_diff)

median(io$mvar_diff)
quantile(io$mvar_diff*100, probs = c(0.25, 0.75))

median(io$tvar_diff)
quantile(io$tvar_diff*100, probs = c(0.25, 0.75))

```
Intraclass coefficients for comparisons of test-retest and inter-observer
```{r}
icc(trt[, c("s1mvar", "s2mvar")],"twoway","agreement","single")
icc(trt[, c("s1tvar", "s2tvar")],"twoway","agreement","single")

icc(io[, c("o1mvar", "o2mvar")],"twoway","agreement","single")
icc(io[, c("o1tvar", "o2tvar")],"twoway","agreement","single")

```
Check some max and mins for the plot
```{r}
max(trt$mvar_diff*100)
max(trt$mvar_mean*100)
min(trt$mvar_diff*100)
min(trt$mvar_mean*100)

max(trt$tvar_diff*100)
max(trt$tvar_mean*100)
min(trt$tvar_diff*100)
min(trt$tvar_mean*100)
```

Check some max and mins for the plot
```{r}
max(io$mvar_diff*100)
max(io$mvar_mean*100)
min(io$mvar_diff*100)
min(io$mvar_mean*100)
max(io$tvar_diff*100)
max(io$tvar_mean*100)
min(io$tvar_diff*100)
min(io$tvar_mean*100)
```
Bland-Altman 1986 limits of agreement for test-retest and inter-observer
```{r}
trt_m_blandr <- blandr.statistics(trt$s1mvar, trt$s2mvar, sig.level = 0.95, LoA.mode = 1)
trt_t_blandr <- blandr.statistics(trt$s1tvar, trt$s2tvar, sig.level = 0.95, LoA.mode = 1)
io_m_blandr <- blandr.statistics(io$o1mvar, io$o2mvar, sig.level = 0.95, LoA.mode = 1)
io_t_blandr <- blandr.statistics(io$o1tvar, io$o2tvar, sig.level = 0.95, LoA.mode = 1)

```

Bland-Altman plot for test-rest and inter-observer mitral and tricuspid
```{r}
library(ggplot2)
library(cowplot)

a <- c(14)
b <- c(10)
c <- c(10)

xlim <- c(10,65)
ylim <- c(-12.5,12.5)
offset <- 1.6
xpos <- 59.5
# Plot 1: Test-retest for mitral
p1 <- ggplot(trt, aes(x = mvar_mean*100, y = mvar_diff*100)) +
  geom_point(size=1) +
  geom_hline(yintercept = 0,linetype="dashed") +
  geom_hline(yintercept = trt_m_blandr$bias*100) +
  geom_hline(yintercept = trt_m_blandr$upperLOA*100, color = "red", linetype="dashed") +
  geom_hline(yintercept = trt_m_blandr$lowerLOA*100, color = "red", linetype="dashed") +
  ggtitle("A") +
  ylab("Mitral respiratory variation\nsession 1 minus session 2\n(%-points)") +
  xlab("Mean (%-points)") +
  annotate("text", 
         x = xpos, 
         y = trt_m_blandr$upperLOA * 100 + offset, 
         label = sprintf("+1.96SD %.1f", trt_m_blandr$upperLOA * 100), 
         size = 3, 
         color = "red") + 
   annotate("text", 
         x = xpos, 
         y = trt_m_blandr$lowerLOA * 100 + offset, 
         label = sprintf("-1.96SD %.1f", trt_m_blandr$lowerLOA * 100), 
         size = 3, 
         color = "red") + 
  theme_minimal() +
  theme(panel.grid = element_blank(),
        plot.title = element_text(size=a, face = "bold"),
        axis.title = element_text(size=b),
        axis.text = element_text(size=c),
        axis.line = element_line(color = "black"),  # Adds axis lines
        axis.ticks = element_line(color = "black")) +  # Adds axis ticks
  coord_fixed(ratio = 1) +
  xlim(xlim) +
  ylim(ylim)

# Plot 2: Test-retest for tricuspid
p2 <- ggplot(trt, aes(x = tvar_mean*100, y = tvar_diff*100)) +
  geom_point(size=1) +
  geom_hline(yintercept = 0,linetype="dashed") +
  geom_hline(yintercept = trt_t_blandr$bias*100) +
  geom_hline(yintercept = trt_t_blandr$upperLOA*100, color = "red", linetype="dashed") +
  geom_hline(yintercept = trt_t_blandr$lowerLOA*100, color = "red", linetype="dashed") +
  ggtitle("B") +
  ylab("Tricuspid respiratory variation\nsession 1 minus session 2\n(%-points)") +
  xlab("Mean (%-points)") +
  annotate("text", 
         x = xpos, 
         y = trt_t_blandr$upperLOA * 100 + offset, 
         label = sprintf("+1.96SD %.1f", trt_t_blandr$upperLOA * 100), 
         size = 3, 
         color = "red") + 
   annotate("text", 
         x = xpos, 
         y = trt_t_blandr$lowerLOA * 100 + offset, 
         label = sprintf("-1.96SD %.1f", trt_t_blandr$lowerLOA * 100), 
         size = 3, 
         color = "red") + 

  theme_minimal() +
  theme(panel.grid = element_blank(),
        plot.title = element_text(size=a, face = "bold"),
        axis.title = element_text(size=b),
        axis.text = element_text(size=c),
        axis.line = element_line(color = "black"),  # Adds axis lines
        axis.ticks = element_line(color = "black")) +  # Adds axis ticks
  coord_fixed(ratio = 1) +
  xlim(xlim) +
  ylim(ylim)

# Plot 3: Inter-observer for mitral
p3 <- ggplot(io, aes(x = mvar_mean*100, y = mvar_diff*100)) +
  geom_point(size=1) +
    geom_hline(yintercept = 0,linetype="dashed") +
  geom_hline(yintercept = io_m_blandr$bias*100) +
  geom_hline(yintercept = io_m_blandr$upperLOA*100, color = "red", linetype="dashed") +
  geom_hline(yintercept = io_m_blandr$lowerLOA*100, color = "red", linetype="dashed") +
  ggtitle("C") +
  ylab("Mitral respiratory variation\nobserver 1 minus observer 2\n(%-points)") +
  xlab("Mean (%-points)") +
  annotate("text", 
         x = xpos, 
         y = io_m_blandr$upperLOA * 100 + offset, 
         label = sprintf("+1.96SD %.1f", io_m_blandr$upperLOA * 100), 
         size = 3, 
         color = "red") + 
   annotate("text", 
         x = xpos, 
         y = io_m_blandr$lowerLOA * 100 + offset, 
         label = sprintf("-1.96SD %.1f", io_m_blandr$lowerLOA * 100), 
         size = 3, 
         color = "red") + 
  
  theme_minimal() +
  theme(panel.grid = element_blank(),
        plot.title = element_text(size=a, face = "bold"),
        axis.title = element_text(size=b),
        axis.text = element_text(size=c),
        axis.line = element_line(color = "black"),  # Adds axis lines
        axis.ticks = element_line(color = "black")) +  # Adds axis ticks
  coord_fixed(ratio = 1) +
  xlim(xlim) +
  ylim(ylim)

# Plot 4: Inter-observer for tricuspid
p4 <- ggplot(io, aes(x = tvar_mean*100, y = tvar_diff*100)) +
  geom_point(size=1) +
    geom_hline(yintercept = 0,linetype="dashed") +
  geom_hline(yintercept = io_t_blandr$bias*100) +
  geom_hline(yintercept = io_t_blandr$upperLOA*100, color = "red", linetype="dashed") +
  geom_hline(yintercept = io_t_blandr$lowerLOA*100, color = "red", linetype="dashed") +
  ggtitle("D") +
  ylab("Tricuspid respiratory variation\nobserver 1 minus observer 2\n(%-points)") +
  xlab("Mean (%-points)") +
  annotate("text", 
         x = xpos, 
         y = io_t_blandr$upperLOA * 100 + offset, 
         label = sprintf("+1.96SD %.1f", io_t_blandr$upperLOA * 100), 
         size = 3, 
         color = "red") + 
   annotate("text", 
         x = xpos, 
         y = io_t_blandr$lowerLOA * 100 + offset, 
         label = sprintf("-1.96SD %.1f", io_t_blandr$lowerLOA * 100), 
         size = 3, 
         color = "red") + 
  theme_minimal() +
  theme(panel.grid = element_blank(),
        plot.title = element_text(size=a, face = "bold"),
        axis.title = element_text(size=b),
        axis.text = element_text(size=c),
        axis.line = element_line(color = "black"),  # Adds axis lines
        axis.ticks = element_line(color = "black")) +  # Adds axis ticks
  coord_fixed(ratio = 1) +
  xlim(xlim) +
  ylim(ylim)

# Arrange the plots in a 2x2 grid
plot_grid(p1, p2, p3, p4, ncol = 2, align = "v")

```



